The Need to Develop a Holistic Infectious Disease Surveillance System

### 2024-10-30 manuscript/manuscript.qmd
Update references in intro

In this paper we examine how the use of imperfect diagnostic tests affects the performance of outbreak detection for measles outbreaks in the context of a febrile rash surveillance system that includes both measles and non-measles cases.
Because measles symptoms are non-specific, it is important to account for non-measles sources of febrile rash e.g., rubella, parvovirus, varicella, etc., producing the potential for false positive results in the context of imperfect tests.
Currently, measles outbreaks are declared on the basis of either suspected measles cases (i.e., an individual with fever and maculopapular rash [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018]) alone, cases confirmed by enzyme immunoassay / enzyme-linked immunosorbent assay, from here-on in referred to as ELISA, to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
Each of these detection systems have their flaws.
Although clinical case definition is very fast and requires limited resources, it is highly sensitive, and in the face of high “background noise” from non-measles sources of febrile rash, can lead to low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
And while ELISA confirmation is the standard diagnostic test for measles surveillance and has higher specificity [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], the training and facility requirements generally mean that samples must be transported from the point of care to a separate laboratory, which incurs both costs and delays _**[ADD REFERENCE HERE]**_.
In resource-poor settings these delays may be days to weeks.
A rapid diagnostic test may meet the WHO's definition and requirements for using in a surveillance setting, if not for individual patient care [@worldhealthorganizationSelectionUseEssential2021].
Recent developments show encouraging signs in the field as well as in theory, providing a compromise between diagnostic accuracy and timeliness in most [@warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020], though not all [@seninMeaslesIgMRapid2024], settings.

>>>>

In this paper we examine how the use of imperfect diagnostic tests affects the performance of outbreak detection for measles outbreaks in the context of a febrile rash surveillance system that includes both measles and non-measles cases.
Because measles symptoms are non-specific, it is important to account for non-measles sources of febrile rash e.g., rubella, parvovirus, varicella, etc., producing the potential for false positive results in the context of imperfect tests.
Currently, measles outbreaks are declared on the basis of either suspected measles cases (i.e., an individual with fever and maculopapular rash [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018]) alone, cases confirmed by enzyme immunoassay / enzyme-linked immunosorbent assay, from here-on in referred to as ELISA, to detect the presence of measles-specific IgM antibodies, or a combination of the two, depending on the elimination status of the region, with countries nearer elimination increasing the use of ELISA diagnostic confirmation for suspected cases [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018].
Each of these detection systems have their flaws.
Although clinical case definition is very fast and requires limited resources, it is highly sensitive, and in the face of high “background noise” from non-measles sources of febrile rash, can lead to low positive predictive value (PPV), i.e., the probability that an alert accurately captures an outbreak [@hutchinsEvaluationMeaslesClinical2004].
And while ELISA confirmation is the standard diagnostic test for measles surveillance and has higher specificity [@essentialprogrammeonimmunizationepiimmunizationvaccinesandbiologicalsivbMeasles2018], the training and facility requirements generally mean that samples must be transported from the point of care to a separate laboratory, which incurs both costs and delays [@essentialprogrammeonimmunizationepiimmunizationClinicalSpecimensLaboratory2018; @worldhealthorganizationMeaslesOutbreakGuide2022; @brownRapidDiagnosticTests2020]
In resource-poor settings these delays may be days to weeks.
A rapid diagnostic test may meet the WHO's definition and requirements for using in a surveillance setting, if not for individual patient care [@worldhealthorganizationSelectionUseEssential2021].
Recent developments show encouraging signs in the field as well as in theory, providing a compromise between diagnostic accuracy and timeliness in most [@shonhaiInvestigationMeaslesOutbreak2015; @warrenerEvaluationRapidDiagnostic2023; @brownRapidDiagnosticTests2020], though not all [@seninMeaslesIgMRapid2024], settings.
